---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# fstpindia

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)
[![R-CMD-check](https://github.com/agentsforsci-ghe/fstp-eval-india/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/agentsforsci-ghe/fstp-eval-india/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The goal of fstpindia is to provide the data of the report *Evaluation of
FSTPs and STP Co-treatment Systems across India* by the Centre for
Science and Environment (CSE), New Delhi, 2023, as tidy datasets in R
and as CSV and XLSX files. The report evaluated 69 faecal sludge
treatment plants (FSTPs) and sewage treatment plants (STPs) with
co-treatment of faecal sludge in eight Indian states.

## Installation

You can install the development version of fstpindia from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("agentsforsci-ghe/fstp-eval-india")
```

```{r}
## Run the following code in console if you don't have the packages
## install.packages(c("dplyr", "knitr", "readr", "stringr", "kableExtra"))
library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(kableExtra)
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

1.  Click Download CSV. A window opens that displays the CSV in
    your browser.
2.  Right-click anywhere inside the window and select "Save Page As...".
3.  Save the file in a folder of your choice.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

extdata_path <- "https://github.com/agentsforsci-ghe/fstp-eval-india/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  dplyr::mutate(file_name = str_remove(file_name, ".rda")) |> 
  dplyr::rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

## Data

The package provides access to ...

```{r}
library(fstpindia)
```

### bod_compliance

The dataset `bod_compliance` contains data about ... It has
`r nrow(bod_compliance)` observations and `r ncol(bod_compliance)`
variables

```{r}
bod_compliance |> 
  head(3) |> 
  knitr::kable()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "bod_compliance.rda") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Example

Describe here what the plot below shows and why it is a useful first look
at the data.

```{r}
library(fstpindia)
# install.packages("ggplot2")
# library(ggplot2)

# A first plot of the data: replace the aesthetics with variables from
# bod_compliance, then uncomment the block and the library call above.
# bod_compliance |>
#   ggplot(aes(x = , y = )) +
#   geom_point() +
#   labs(x = "", y = "", title = "")
```

## License

Data are available as
[CC-BY](https://github.com/agentsforsci-ghe/fstp-eval-india/blob/main/LICENSE.md).

## Citation

Please cite this package using:

```{r}
citation("fstpindia")
```
